The Astropy Project:

### 2017-09-29 main.tex
python

\begin{abstract}
% I (Adrian) took a first stab at the abstract, but it needs some work. Feel
% free to modify!
The \astropy\ project supports and fosters the development of open-source Python
packages that provide commonly-needed functionality to the astronomical
community.
A key element of the project is the \astropy\ core package, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy\
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate [...]
We conclude with a future outlook of planned new features and directions for the
broader \astropy\ ecosystem.
\end{abstract}

@@

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing fits files is one example where it is time
consuming to implement the fits standard correctly and
programming is much faster when using an existing library perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy\ project aims to provide a
library of core astronomical functionality in the Python programming
language.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
%
% - reference APE2 for versioning and APE10 for dropping python2 support
% - shall we maybe refer to the release calendar, too?
% - should we mention that v3.0 is the exception to the version numbering
%   scheme we use by not being 2.1?
\par The general scheme for the releases and version numbering that the core
package has adopted if the following. \astropypkg\ has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features. The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports Python2, and will receive bug fixes until the
end of 2019.
%
\par The version numbering of \astropypkg\ reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advances
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

\begin{figure}
\includegraphics[width=\textwidth]{convolution_example.png}
An example showing different modes of convolution available in the \python ecosystem.  The red x's mark pixels that are set to NaN in the original data (a).  If the data are convolved with a Gaussian kernel on a 9x9 grid using scipy's direct convolution (b), any pixel within range of the original NaN pixels is also set to NaN.  Panel (c) shows what happens if the NaNs are set to zero first: the originally NaN regions are depressed relative to their surroundings.  Finally, panel (d) shows \astropypkg's convolution behavior, where the missing pixels are replaced with values interpolated from their surroundings using the convolution kernel.
\end{figure}

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
% - should we use links to the appropriate github repo?
%
\par In addition to astronomy related packages and libraries, the \astropy\
Project also contains and supports several infrastructure package that are
very generic. The following sections describes the most widely used
selection of them.
%
\subsubsection{Package template}
% Brigitta may write this up if there is no other takers
\par \astropy\ provides a template package that any Python package is welcome to
use (many affiliated packages do so). The template contains a ready-to-go
package layout. It provides infrastructure such as documentation tools,
testing framework and CI templates and configurations, Cython integration,
and documentation on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
\texttt{ci-helpers} is a collection of scripts that empowers package
maintainers to control their testing set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy\ ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for TravisCI and AppveyorCI.
%
\subsection{sphinx extensions}
% Tom R

@@

dropping python2 support, growths of affiliated packages
Summary

>>>>

\begin{abstract}
% I (Adrian) took a first stab at the abstract, but it needs some work. Feel
% free to modify!
The \astropy\ project supports and fosters the development of open-source
\python\ packages that provide commonly-needed functionality to the astronomical
community.
A key element of the project is the \astropy\ core package, which serves as the
foundation for more specialized projects and packages.
In this article, we provide an overview of the organization of the \astropy\
project and summarize key features in the core package as of the recent major
release, version 2.0.
We then describe the project infrastructure designed to facilitate [...]
We conclude with a future outlook of planned new features and directions for the
broader \astropy\ ecosystem.
\end{abstract}

@@

The implementation of astronomical software, be it for a package meant
for wide distribution or scripts and programs for a specific research
project, can be eased by the use of a library that provides core
functionality that is common to many astronomical tasks. Reading and
writing fits files is one example where it is time
consuming to implement the fits standard correctly and
programming is much faster when using an existing library perform this
task. Another example of such a common task is dealing with different
astronomical coordinate systems. The \astropy\ project aims to provide a
library of core astronomical functionality in the \python\ programming
language.

@@

\subsection{Release cycle and Long Term Support}
% This could also include information about the development cycle
% Brigitta can writes this up if there's no other taker
%
% - reference APE2 for versioning and APE10 for dropping python2 support
% - shall we maybe refer to the release calendar, too?
% - should we mention that v3.0 is the exception to the version numbering
%   scheme we use by not being 2.1?
\par The general scheme for the releases and version numbering that the core
package has adopted if the following. \astropypkg\ has new feature releases every
6 months, and between feature releases additional bugfix releases that
contain only bugfixes but no new features. The timing of
%
\par Long-term support (LTS) releases continue to receive bugfixes for 2
years with no changes to the API. They are ideal for pipelines and other
applications where API stability is essential. The latest LTS release is
also the last one that supports \python 2, and will receive bug fixes until the
end of 2019.
%
\par The version numbering of \astropypkg\ reflects on the release scheme
described above. The core package uses the form x.y.z, where x is advances
on LTS, y is advanced for a non-LTS feature release, and z is advanced on
bugfix releases.

@@

\begin{figure}
\includegraphics[width=\textwidth]{convolution_example.png}
An example showing different modes of convolution available in the \python\ ecosystem.  The red x's mark pixels that are set to NaN in the original data (a).  If the data are convolved with a Gaussian kernel on a 9x9 grid using scipy's direct convolution (b), any pixel within range of the original NaN pixels is also set to NaN.  Panel (c) shows what happens if the NaNs are set to zero first: the originally NaN regions are depressed relative to their surroundings.  Finally, panel (d) shows \astropypkg's convolution behavior, where the missing pixels are replaced with values interpolated from their surroundings using the convolution kernel.
\end{figure}

@@

\section{Infrastructure for affiliated packages}
\label{sec:infrastructure}
% - should we use links to the appropriate github repo?
%
\par In addition to astronomy related packages and libraries, the \astropy\
Project also contains and supports several infrastructure package that are
very generic. The following sections describes the most widely used
selection of them.
%
\subsubsection{Package template}
% Brigitta may write this up if there is no other takers
\par \astropy\ provides a template package that any \python\ package is welcome to
use (many affiliated packages do so). The template contains a ready-to-go
package layout. It provides infrastructure such as documentation tools,
testing framework and CI templates and configurations, Cython integration,
and documentation on how to make it all work.
%
\subsection{Continuous integration helpers}
% Brigitta
\texttt{ci-helpers} is a collection of scripts that empowers package
maintainers to control their testing set up and installation process for
various continuous integration services via a set of environment
variables. While the current development is mostly driven by the needs of
the \astropy\ ecosystem, the actual usage of this package is extremely
widespread. Currently we support setups for TravisCI and AppveyorCI.
%
\subsection{sphinx extensions}
% Tom R

@@

dropping \python 2 support, growths of affiliated packages
Summary
